Loop-Aware Optimizations in PyPy's Tracing JIT

### 2011-06-27 talk/iwtc11/paper.tex
more related work

\subsection{Allocation Removals}
PyPy's allocation removal optimization \cite{bolz_allocation_2011} makes it
possible to identify objects that are allocated within the loop but never
escape it. Those objects have to be allocated in the loop, but no outside
object ever gets a reference short lived objects with no references outside the
loop. This
is performed by processing the operations in order and
optimistically removing every \lstinline{new} operation. Later on if
it is discovered that a reference to the object escapes the loop, the
\lstinline{new} operation is inserted at this point. All operations
(\lstinline{get}, \lstinline{set} and \lstinline{guard}) on the removed objects
are also removed and the optimizer needs to keep track of the value of all used
attributes of the object.

@@

SPUR \cite{bebenita_spur:_2010} implements loop-invariant code motion
directly, by explicitly marking as loop-invariant all variables that stay the
same along all looping paths and then moving all pure computation that depends
only on these variables out of the loop. SPUR can also hoist loads out of the
loop if nothing in the loop can ever write to the memory location. It can also
move allocations out of the loop, but does not replace the object by its fields.
This saves only the allocation, not the access to the object fields.

XXX
% section Related Work (end)

@@

\acks

Acknowledgments, if needed.

>>>>

\subsection{Allocation Removals}
\label{sub:allocation}

PyPy's allocation removal optimization \cite{bolz_allocation_2011} makes it
possible to identify objects that are allocated within the loop but never
escape it. Those objects have to be allocated in the loop, but no outside
object ever gets a reference short lived objects with no references outside the
loop. This
is performed by processing the operations in order and
optimistically removing every \lstinline{new} operation. Later on if
it is discovered that a reference to the object escapes the loop, the
\lstinline{new} operation is inserted at this point. All operations
(\lstinline{get}, \lstinline{set} and \lstinline{guard}) on the removed objects
are also removed and the optimizer needs to keep track of the value of all used
attributes of the object.

@@

Both the Hotpath VM \cite{gal_hotpathvm:_2006} and SPUR
\cite{bebenita_spur:_2010} implements loop-invariant code motion
directly, by explicitly marking as loop-invariant all variables that stay the
same along all looping paths and then moving all pure computation that depends
only on these variables out of the loop. SPUR can also hoist loads out of the
loop if nothing in the loop can ever write to the memory location. It can also
move allocations out of the loop, but does not replace the object by its fields.
This saves only the allocation, not the access to the object fields.

The type specialization described by Gal \etal \cite{gal_trace-based_2009} can
be seen as doing a similar optimization (again by manually implementing it)
than the one described in Section~\ref{sub:allocation}: The effect of both is
that type checks are fully done before a loop is even entered.

% section Related Work (end)

@@

%\acks
%Acknowledgments, if needed.
